---
title: "DATA SCIENCE AND ANALYTICS"
author: "Ayşe Ceren Çiçek"
output: 
  pdf_document: default
  html_document: default
  word_document: default
subtitle: Week 2
font-family: Gill Sans
---

# Data Transformation

Most of the code examples written down below, taken from this website https://r4ds.had.co.nz/transform.html

<script type="text/javascript">
  // When the document is fully rendered...
  $(document).ready(function() {
    // ...select all header elements...
    $('h1, h2, h3, h4, h5').each(function() {
      // ...and add an id to them corresponding to their 'titles'
      $(this).attr('id', $(this).html());
    });
  });
</script>

## TABLE OF CONTENTS


- <a href="#Import libraries">Importing Libraries</a><br>
- <a href="#Dataset">Dataset Informations</a><br>
- <a href="#dplyr Basics">dplyr Basics</a><br>
  - <a href="#Filter">Filter</a><br>
    - <a href="#Logical Operators">Logical Operators</a><br>
    - <a href="#Missing Values">Missing Values</a><br>
  - <a href="#Arrange">Arrange</a><br>
  - <a href="#Select">Select</a><br>
  - <a href="#Mutate">Mutate</a><br>
  - <a href="#Summarise">Summarise</a><br>


## Import libraries

```{r, warning=FALSE,message=FALSE,error=FALSE, results='hide'}
#install.packages("nycflights13")
library(nycflights13)
library(tidyverse)
```
## Dataset

The dataset contains data for all flights that departed NYC (i.e. JFK, LGA or EWR) in 2013. Column informations are listed below:

- year, month, day: Date of departure.

- dep_time, arr_time: Actual departure and arrival times (format HHMM or HMM), local tz.

- sched_dep_time, sched_arr_time: Scheduled departure and arrival times (format HHMM or HMM), local tz.

- dep_delay, arr_delay: Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.

- carrier: Two letter carrier abbreviation. See airlines to get name.

- flight: Flight number.

- tailnum: Plane tail number. See planes for additional metadata.

- origin, dest: Origin and destination. See airports for additional metadata.

- air_time: Amount of time spent in the air, in minutes.

- distance: Distance between airports, in miles.

- hour, minute: Time of scheduled departure broken into hour and minutes.

- time_hour: Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.

```{r}
flights
```

To see the whole dataset, you can run View(flights).

```{r}
summary(flights)
```


## dplyr Basics 

- Pick observations by their values (filter()).
- Reorder the rows (arrange()).
- Pick variables by their names (select()).
- Create new variables with functions of existing variables (mutate()).
- Collapse many values down to a single summary (summa rize()).


### Filter

filter() allows you to subset observations based on their values.

We can select all flights on May 1st with.

```{r}
filter(flights, month == 5, day == 1)
```

```{r}
(june25 <- filter(flights, month == 6, day == 25))
```

#### Logical Operators

```{r}
filter(flights, month == 6 | month == 12)
```

x %in% y --> this will select every row where x is one of the values in y

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
nov_dec
```


```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
```

#### Missing Values

```{r}
x <- NA
is.na(x)
```

```{r}
df <- tibble(x = c(1, NA, 3))

filter(df, is.na(x) | x > 1)
```

### Arrange

arrange() takes a data frame and a set of column names to order by.

```{r}
arrange(flights, year, month, day)
```

desc() to re-order by a column in descending order.

```{r}
arrange(flights, desc(dep_delay))
```


### Select

select() allows you to rapidly zoom in on a useful subset using operations based on the names of the variables.

```{r}
select(flights, year, month, day)
```


```{r}
#select all columns between year and day.
select(flights, year:day)
```


```{r}
#select all columns except those from year to day
select(flights, -(year:day))
```



```{r}
#everything() moves some variables to the start of the data frame
select(flights, time_hour, air_time, day, everything())
```


### Mutate

mutate() always adds new columns at the end of the dataset.

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

If you only want to keep the new variables, use transmute().

```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```


```{r}
(x <- 1:10)
lag(x)
lead(x)
```


```{r}
cumsum(x)
#cumulative means
cummean(x)
```

### Summarise

summarise() collapses a data frame to a single row.

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r, warning=FALSE, error=FALSE, message=FALSE}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```


#### Combining multiple operations with the pipe(%>%)

```{r, message=FALSE}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

- Counts
- Useful Summary Functions
- Grouping by multiple variables
- Ungrouping

