---
title: "CENG 4515 DATA SCIENCE AND ANALYTICS"
author: "Ayşe Ceren Çiçek"
output: 
  html_document: default
  pdf_document: default
  word_document: default
subtitle: HOMEWORK - 2
font-family: Gill Sans
---

Dataset: https://www.kaggle.com/uciml/breast-cancer-wisconsin-data

The columns are:

- id: ID number

- diagnosis: The diagnosis of breast tissues (M = malignant, B = benign)

- radius_mean: mean of distances from center to points on the perimeter

- texture_mean: standard deviation of gray-scale values

- perimeter_mean: mean size of the core tumor

And the other columns are: area_mean, smoothness_mean, compactness_mean, concavity_mean, concave points_mean, symmetry_mean, fractal_dimension_mean, radius_se, texture_se, perimeter_se, area_se, smoothness_se, compactness_se, concavity_se, concave points_se, symmetry_se, fractal_dimension_se, radius_worst, texture_worst, perimeter_worst, area_worst, smoothness_worst, compactness_worst, concavity_worst, concave points_worst, symmetry_worst, fractal_dimension_worst. 



## Importing libraries

```{r, warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(tidyverse)
library(plyr)
library(ggplot2)
library(gplots)
library(caret)
library(ROCR)
library(logistf)
library(PerformanceAnalytics)
library(GGally)
library(corrplot)
library(pheatmap)
library(factoextra)
library(ggfortify)
library(sparklyr)
library(pipeR)
```


## Loading dataset

```{r}
# Load dataset
data <- read.csv("data.csv")

# Show first 5 columns of the data
head(data, n=5)
```



# Data preprocessing

Let's summarize our data.

```{r}
summary(data)
```

The original dataset does not contain the X columns we will remove it.

```{r}
data <- subset(data, select=-X)
```


```{r}
# Check null values
which(is.na(data))
```

We will remove id column and replace diagnosis values B with 0 and M with 1.

```{r}
data$id <- NULL
data$diagnosis <- ifelse(data$diagnosis == "B", 0, 1)
```

Let's check column types.

```{r}
sapply(data, class)
```
All of the columns seems numeric so we can draw a heatmap.

```{r}
heatmap(cor(data[,-1]))
```

```{r}
data_corr = cbind(diagnosis=data[,c(1)], data[,c(22:31)])
```



```{r, warning=FALSE}
# Draw correlation matrix 
ggcorr(data, nbreaks=8, label=TRUE, label_size=2, size = 1.8, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45) + ggtitle("Breast Cancer Correlation Matrix")
corr_data = cor(data)

corr_data = round(corr_data, 4)

# Remove low relationship
corr_data[abs(corr_data) < 0.7] = NA 

# Drop duplicates and meaningless information
corr_data[lower.tri(corr_data, diag = TRUE)] = NA  

# Turn it into a 3-column table
corr_data=as.data.frame(as.table(corr_data))  
corr_data=na.omit(corr_data) 

#Sort by highest correlation
corr_data=corr_data[order(-abs(corr_data$Freq)),]    
corr_data
```

## Building a logistic regression model (Base model)

```{r}
set.seed(123)

# Train test split
splitted<-createDataPartition(data$diagnosis, p=0.8, list=FALSE)

# Train data
X_train<-data[splitted,-1]
y_train<-data$diagnosis[splitted] 

# Test data
X_test<-data[-splitted,-1]
y_test<-data$diagnosis[-splitted]
```


```{r, warning=FALSE}
# Logistic regression model with all variables
model <- glm(formula = y_train ~ ., family = binomial, data = X_train)
summary(model)
```


```{r}
# Subset without correlated variables
subset.data<-subset(data, select=-c(area_mean,radius_mean,area_worst,compactness_mean,perimeter_worst,compactness_se,concavity_worst,fractal_dimension_worst))
```


```{r}
# Train data
X_train<-subset.data[splitted,-1]
y_train<-subset.data$diagnosis[splitted]

# Test data
X_test<-subset.data[-splitted,-1]
y_test<-subset.data$diagnosis[-splitted]
```


```{r, warning=FALSE}
# Logistic regression model without correlated variables
model <- glm(formula = y_train ~ ., family = binomial, data = X_train)
summary(model)
```


```{r, warning=FALSE}
anova(model)
```



```{r}
# Make predictions
prediction <- ifelse(predict(object=model, newdata=X_test, type = "response") > 0.5, 1, 0)

# Generate confusion matrix
conf.matrix <- confusionMatrix(table(prediction, y_test), positive="1")
conf.matrix
```

```{r}
# Heatmap visualization of confusion matrix
table <- data.frame(conf.matrix$table)

plotTable <- table %>%
  group_by(y_test) %>%
  mutate(prop = Freq/sum(Freq))

ggplot(data =  plotTable, mapping = aes(x = y_test, y = prediction, alpha = prop)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1, color="white", size=10) +
  scale_fill_gradient(low = "blue", high = "navyblue") +
  theme_bw() + theme(legend.position = "none")

```


## Applying PCA

```{r,  results='hide'}
# Feature scaling
scale(X_train)
scale(X_test)
```


```{r,}
# Apply PCA
pca <- prcomp(X_train[,2:ncol(X_train)],center = TRUE, scale=TRUE)
pca_test<-prcomp(X_test[,2:ncol(X_test)],center = TRUE, scale=TRUE)

# Plot PCA
plot(pca, type='l', main="PCA - Principal Components Analysis Chart", col="red")
```

```{r}
#Get eigenvalues
explained.variance <- pca$sdev^2
explained.variance
```


```{r}
# Plot PCA
pca.var <- get_pca_var(pca)

kmean <- kmeans(pca.var$coord, centers = 5, nstart=25)
group <- as.factor(kmean$cluster)

fviz_pca_var(pca, col.var=group, palette='jco', legend.title='Cluster')
```


```{r}
# Plot PCA
autoplot(pca, data = X_train)
```

## Building a logistic regression model on the transformed data


```{r}
# Logistic regression model
model_pca <- glm(y_train ~ .-PC11 -PC6 -PC8 -PC10 -PC5 -PC9 -PC13 -PC14 -PC12 -PC7,family=binomial, data=pca$x[, c(1:14)] %>% data.frame())
summary(model_pca)
```


```{r}
# Make prediction
prediction.pca <- ifelse(predict(model_pca) > 0.5,1,0)

# Generate confusion matrix
conf.matrix.pca <- confusionMatrix(table(prediction.pca, y_train), positive="1")
conf.matrix.pca
```

```{r}
conf.matrix.pca$table
```


```{r}
# Heatmap visualization of the confusion matrix
table.pca <- data.frame(conf.matrix.pca$table)

plotTablePCA <- table.pca %>%
  group_by(y_train) %>%
  mutate(prop = Freq/sum(Freq))

ggplot(data =  plotTablePCA, mapping = aes(x = prediction.pca, y = y_train, alpha = prop)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1, color="white", size=10) +
  scale_fill_gradient(low = "blue", high = "navyblue") +
  theme_bw() + theme(legend.position = "none")

```



## Building a machine learning Pipeline


```{r, warning=FALSE, error=FALSE}
X_train_scaled = scale(X_train)
log.rec.model <- trainControl(method = "glm")
pca.model <- trainControl(method = "prcomp")

#log_reg_model <-  pipeline(X_train_scaled, pca.model, log.rec.model)


#sc <- spark_connect(master = "local", spark_version = "2.2.0")


# pipeline <- ml_pipeline(sc) %>%
#   ft_dplyr_transformer(
#     tbl = pca
#     ) %>% 
#   ml_logistic_regression()
```




